\textbf{Appendix}\\ Bound within Boundaries: How Well Do Protected Areas Match Movement Corridors of Their Most Mobile Protected Species?

### 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-04-17.tex
first commit

% Connectivity Assessment
In recent years, there has been a growing body of literature that uses animal
relocation data to identify dispersal corridors and to assess connectivity on
large scales \citep{Chetkiewicz.2006}. Identification of potential dispersal
corridors between suitable habitat patches traditionally relies on the
estimation of a permeability surface (elsewhere referred to as resistance
surface, when inverted), which renders the ease or willingness at which the
focal species traverses a specific landscape. Latest empirical methods to
predict landscape permeability based on recorded GPS relocation data include
\textit{resource selection functions} (RSF, \cite{Boyce.2002}), \textit{step
selection functions} (SSF, \cite{Fortin.2005}), and \textit{path selection
functions} (PSF, \cite{Cushman.2010}). All methods have in common that they
infer habitat preferences of the focal species by comparing spatial covariates
(e.g. environmental and anthropogenic) at visited locations to the same spatial
covariates at randomly selected locations that are deemed available but
presumably unused by the animal. Based on inferred habitat preferences the
methods further assume that habitats preferred by the focal species provide high
permeability, whereas avoided habitats provide low permeability. Regardless of
the method chosen (i.e. RSF, SSF, or PSF), all approaches require adequate GPS
relocation data that is representative of the process being studied. More
specifically, it has been shown that GPS relocation data collected on dispersing
individuals are better suited for detecting dispersal corridors than GPS
relocation data collected on resident individuals \citep{Elliot.2014}. Few
studies have, however, partitioned their GPS relocation data according to
behavioral modes such as \textit{resident} vs. \textit{dispersing} or
\textit{active} vs. \textit{inactive} periods of movement \citep{Wilson.2012,
Vasudev.2015}. As a consequence, most permeability surfaces upon which dispersal
corridors are identified are created using GPS relocation data that was
collected on resident individuals. This introduces severe biases and
substantially reduces the power to reveal meaningful corridors, for dispersing
individuals have different needs and motivations compared to resident
individuals (\citeauthor{Killeen.2014}, 2014; \citeauthor{Elliot.2014}, 2014;
\citeauthor{Cozzi.2020}, unpublished). For instance, \cite{Elliot.2014}
demonstrated marked differences in predicted permeability surfaces for
dispersing lions in the KAZA-TFCA depending on whether permeability was assessed
using data collected during residence or dispersal. \cite{Jackson.2016} made
similar observations for African wild dogs and revealed higher permeability for
dispersers than for residents. Moreover, \cite{Abrahms.2017} found a
significantly greater overlap between predicted and observed corridors for three
dispersing wild dogs when predictions were based on information collected only
during wild dog activity, rather than across the entire day.

@@

% Large Study Area
The \textit{extended study area} (centered at -17\degree 13'9''S, 23\degree
56'4''E; elevation ca. 1,030 m, red rectangle in \Cref{StudyArea}a) was confined
by a bounding box encompassing the entire KAZA-TFCA. The KAZA-TFCA lies in the
basins of the Okavango and Zambezi rivers and embraces parts of Angola,
Botswana, Namibia, Zimbabwe, and Zambia. With a total area of over 520,000
km\textsuperscript{2}, it contains several already-existing national parks (NPs)
and other protected areas. It constitutes the earth's largest transboundary
conservation area, which resulted in a rectangular study area of over 1.3 Mio
km\textsuperscript{2}. Similar to the core study area, the extended study area
hosts various landscapes, including savanna, grassland, and dry or moist
woodland habitats. Despite 36 NPs and other protected areas, there is
considerable human influence in some regions of the KAZA-TFCA originating from
farms, high human density, and road traffic.

@@

% Water Layers
Water included open waters, rivers, wetlands, and swamps, whereas anything else
was presumed to be dryland. Because the flood extent of the Okavango Delta is
highly variable within and between years, we created weekly (8 days)
``floodmaps'', following the algorithm developed in \cite{Wolski.2017}. The
algorithm uses thresholding of SWIR reflectances of MODIS Terra satellite
imagery (MCD43A4, \cite{Schaaf.2015}) to separate water and dryland. As we
rewrote and slightly amended the original Python code in R, we describe the
underlying principles and indicate the discrepancies to the original algorithm
in \Cref{Appendix:FloodmappingAlgorithm}. While water in the Okavango Delta was
mapped dynamically, we used a static watermap for its surroundings. This static
map was based on Globeland's land cover dataset \citep{Chen.2015}, from which we
only retained the categories \textit{wetland} and \textit{water bodies} and
collectively reclassified them to \textit{water}. Globeland had an original
resolution of 30m x 30m, so we coarsened the layer to 250m x 250m using the mode
of each 250m x 250m cell. Lastly, we improved river representation using the
MERIT Hydro dataset \citep{Yamazaki.2019}, from which we added all rivers with a
width of over 10m to our floodmaps.

% Vegetation Layers
The three vegetation layers that we used in addition to the floodmaps were
derived from the MODIS Terra Vegetation Continuous Fields dataset (MOD44B,
\cite{Dimiceli.2015}). Combined, the three layers added up to 100\% and depicted
the percentage cover of tree-vegetation (henceforth trees), non-tree-vegetation
(henceforth shrubs/grassland), and non-vegetated (henceforth bare land). To
ensure that areas covered by the flood would get 0\% vegetation cover (i.e.
100\% bare land cover), we used our floodmap that aligned with the creation date
of the layers (30 March 2017) and manually defined pixels covered by the flood
as being 100\% bare land. The MODIS vegetation layers had a resolution of 250m x
250m, so we did not need to coarsen or interpolate them.

@@

\subsection{Habitat Selection Model}
\label{Modeling}
% Explanation of iSSF and Extraction
We used an integrated step selection function approach (iSSF, \cite{Avgar.2016})
to investigate dispersers' selection or avoidance for the above mentioned
spatial covariates. In the iSSF framework, covariates experienced along realized
steps are contrasted with covariates experienced along alternative random steps
that the animal could have taken, but decided not to. A step is here defined as
the connecting line between two consecutive GPS relocations
\citep{Turchin.1998}. In contrast to regular SSFs, iSSFs require to include
movement metrics as covariates in the corresponding conditional logistic
regression model. Their inclusion, in turn, allows simultaneous inference on
habitat and movement preferences, as well as to reduce potential biases in
estimated habitat preferences \citep{Forester.2009, Warton.2013, Avgar.2016}.

To conduct an iSSF analysis, we followed recommendations described in
Appendix S1 of \cite{Avgar.2016}. We prepared our GPS relocation data for
iSSF-analysis using the R-package \textit{amt} \citep{Amt.2019} and coerced
relocations recorded during dispersal to steps that were regularly spaced four
hours apart (allowing for a minor mismatch of up to 15 minutes). Steps that were
separated by more than four hours (e.g. due to GPS failure) were omitted from
further analysis. Each remaining step was paired with 24 random steps, generated
by sampling turning angles from a uniform distribution U(\(-\pi, \pi\)) and step
lengths from a gamma distribution that was fitted using realized step lengths
(shape = 0.3677, scale = 6,302)). Together, a realized and its 24 associated
random steps formed a stratum of 25 steps that received a unique identifier.

We extracted spatial covariates along realized and random steps as listed in
\Cref{ExtractedCovars}. For continuous covariates, we calculated the average
value, for binary covariates the percentage cover along the step. Since our
floodmaps were dynamic, we extracted values from the floodmap that was closest
in date to the actual step. We further derived a binary variable indicating
whether or not a step crossed a road and we identified the average distance of
each step to the nearest source of water or the nearest road. We square-rooted
the values for \textit{DistanceToRoads} and \textit{DistanceToWater} to render a
decreasing marginal impact of distance. We standardized extracted covariates to
a mean of zero and a standard deviation of one using a z-score transformation.
Furthermore, we screened covariates for correlation using Pearson's Correlation
Coefficient. None of the covariates were overly correlated (\(|r| > 0.6\),
\cite{Latham.2011}) and we retained all of them for modeling. To run a proper
iSSF analysis, we also included two movement metrics, namely the cosine of the
turning angle (\(cos(ta)\)) and the logarithm of the step length (\(log(sl)\)),
as covariates in our models \citep{Avgar.2016}. The movement metric \(cos(ta\))
served to describe the directionality of a step, as it transformed the circular
measure of (\(-\pi, +\pi\)) into a linear measure (-1, 1). Thus, positive values
indicated forward movements, whereas negative values indicated backward
movements \citep{Turchin.1998}. The movement metric \(log(sl\)), on the other
hand, was included as an indicator of the preferred step length. Since steps
were regularly spaced by four hours, \(log(sl\)) can also be interpreted as
movement rate.

@@

% Mixed Effects Conditional Logistic Regression Model
To deal with multiple individuals, we applied mixed effects conditional logistic
regression analysis following \cite{Muff.2019}. Their approach allowed to
include individuals for whom some covariates did not vary, as well as to model
random intercepts and slopes for each covariate. We implemented their method
using the R-package \textit{glmmTMB} \citep{Mollie.2017}, where we fixed the
variance of the random intercept to an artificially high value of
10\textsuperscript{12} (for an explanation, see \cite{Muff.2019}) and used
animal ID to model random intercepts and slopes.

% AIC Model Selection
We defined the movement metrics \(cos(ta)\) and \(log(sl)\) as core covariates
and ran forward model selection based on Akaike's Information Criterion (AIC,
\cite{Burnham.2002}) for all other spatial covariates. That is, we started with
a basic model and iteratively increased model complexity upon reaching a
full model. We then ranked models according to AIC, assessed relative model
weights, and identified the most parsimonious model. Due to convergence issues,
we were unable to model interactions between spatial covariates.

@@

\subsection{Least-Cost Paths \& Least-Cost Corridors}
To identify movement corridors within the extent of KAZA-TFCA, we calculated
least-cost paths (LCPs) and least-cost corridors (LCCs) between selected points
located in protected areas of our extended study area \citep{Adriaensen.2003,
Sawyer.2011}. Selected points were generated by overlaying the study area with a
regular grid of points, equally spaced 100 km apart and located within protected
areas that contiguously covered at least 700 km\textsuperscript{2}. The
threshold of 700 km\textsuperscript{2} was based on home-range requirements of
African wild dogs reported in \cite{Pomilia.2015} and served to eliminate
locations that were assumed unsuitable for supporting viable wild dog
populations. Finally, we defined centroids as source points for those protected
areas larger than 700 km\textsuperscript{2} that did not get any source points
from the regular grid. In total, we generated 68 points, which resulted in 2,278
unique pairwise combinations and therefore 2,278 unique LCPs and LCCs.

@@

\newpage
\section{Acknowledgements}
All of this work would not have been possible without the aid of several helpful
people. I would first like to thank Prof. Arpat Ozgul, the leader and mastermind
of our research group, whose expertise in statistics, biological processes and
scientific writing was invaluable for my work. Moreover, I want to thank
Gabriele Cozzi for his excellent guidance, all the interesting discussions, his
constructive criticism, valuable revisions in the manuscript, and of course his
constant supply of freshly brewed coffee. Particular gratitude goes to Dominik
Behr, who supervised my work daily and greatly contributed to all aspects of the
thesis. I genuinely enjoyed the frequent meetings and brainstorms and I feel
privileged to have had the opportunity of working with such a great scientist. I
am also greatly indebted to Prof. John Fieberg, who visited our department this
summer and invested a great amount of time to consult and improve all
statistical aspects of the thesis, as well as to deepen my understanding of
movement ecology. Furthermore, I thank the Okavango Research Institute,
especially Piotr Wolski, for sharing all of their Python scripts, floodmaps, and
knowledge on developing dynamic floodmaps. Also, I want to thank Stefan Sommer
for sharing and teaching his hard-earned knowledge on scientific writing.
Additionally, I want to thank all of my colleague master students who have been
wonderful companions and made the time at university a blast. Likewise, I thank
my parents for supporting all of my undertakings and for providing first-class
shelter and food. Last, but most definitely not least, I also greatly thank my
girlfriend, who was always there for me and provided welcome distractions for my
mind.

@@

\newpage
\appendix
\section{Appendix}
\subsection{Floodmapping Algorithm}
\label{Appendix:FloodmappingAlgorithm}
To implement the floodmapping algorithm, we defined two sets of polygons located
in the region of the Okavango Delta (see \Cref{Floodmapping}). The first set
consisted of areas known to be permanent dryland, whereas the second set
consisted of permanent waters. Since we were unable to retrieve the original
polygons used in \cite{Wolski.2017}, we geo-referenced one of their
illustrations and traced the polygons there. After recreation of the polygons,
we used the R-package \textit{getSpatialData} \citep{Schwalb.2018} to download
and pre-process all relatively cloud-free MODIS Terra images (MCD43A4,
\cite{Schaaf.2015}) available for the period of our dispersal events. Assessment
of cloud cover was based on visual inspection of MODIS images on ORI's website
\citep{ORI.2019}. The MCD43A4 MODIS product is particularly useful for dynamic
floodmapping because it is updated with new composite images every 8th day
\citep{Wolski.2017}. After download, we classified each MODIS image into a
binary map of water (flood) and dryland using a threshold that was identified as
follows. First, we extracted all reflectance values of MODIS Terra Band 7 within
the water- and dryland-polygons. Second, we computed histograms of
water-reflectances and dryland-reflectances and empirically verified that
reflectances of the two groups were sufficiently distinct. More specifically, we
checked if superimposing the two histograms resulted in a bimodal plot. This was
said to be achieved if the 99\textsuperscript{th} percentile of
water-reflectances did not severely exceed the 1\textsuperscript{st} percentile
of dryland-reflectances (\(p_{0.99, water} - \frac{10}{255} < p_{0.01,
dryland}\)). Third, if peaks were sufficiently different, we calculated a
threshold (\(t\)) using \Cref{EQ1}:

@@

\newpage
\subsection{Movement Model \& Dispersal Simulation}
\label{Appendix:DispersalSimulation}
\subsubsection{Methods}
To gain more detailed insights into the movement behavior of dispersing wild
dogs and to simulate dispersal events, we fitted a more complex movement model.
The movement model still contained the most parsimonious habitat selection
model, but in addition included interaction terms between movement metrics (i.e.
\(cos(ta)\), \(log(sl\))) and environmental covariates (e.g. water). For
instance, for the environmental covariate \textit{Water}, we created the
interaction terms \textit{cos(ta):Water} and \textit{log(sl):Water}. These
interactions served to learn how dispersers moved differently, depending on
environmental covariates (for comparable work see \cite{Prokopenko.2017}). We
applied the same modeling framework as explained in \Cref{Modeling} and ran
forward model selection based on AIC to identify the most parsimonious movement
model. That is, we departed from the habitat selection model and iteratively
added interactions with movement metrics. Note that we applied forward model
selection only on the newly identified interactions, but not for the terms that
were already in the habitat selection model. Since we attempted to keep our
simulation simple, we did not average models with positive AIC-weights but
simply identified the most parsimonious movement model.

>>>>

% Connectivity Assessment
In recent years, there has been a growing body of literature that uses animal
relocation data to identify dispersal corridors and to assess connectivity on
large scales \citep{Chetkiewicz.2006}. Identification of potential dispersal
corridors between suitable habitat patches traditionally relies on the
estimation of a permeability surface (elsewhere referred to as resistance
surface, when inverted), which renders the ease or willingness at which the
focal species traverses a specific landscape. Latest empirical methods to
predict landscape permeability based on recorded GPS relocation data include
\textit{resource selection functions} (RSF, \citealp{Boyce.2002}), \textit{step
selection functions} (SSF, \citealp{Fortin.2005}), and \textit{path selection
functions} (PSF, \citealp{Cushman.2010}). All methods have in common that they
infer habitat preferences of the focal species by comparing spatial covariates
(e.g. environmental and anthropogenic) at visited locations to the same spatial
covariates at randomly selected locations that are deemed available but
presumably unused by the animal. Based on inferred habitat preferences the
methods further assume that habitats preferred by the focal species provide high
permeability, whereas avoided habitats provide low permeability. Regardless of
the method chosen (i.e. RSF, SSF, or PSF), all approaches require adequate GPS
relocation data that is representative of the process being studied. More
specifically, it has been shown that GPS relocation data collected on dispersing
individuals are better suited for detecting dispersal corridors than GPS
relocation data collected on resident individuals \citep{Elliot.2014}. Few
studies have, however, partitioned their GPS relocation data according to
behavioral modes such as \textit{resident} vs. \textit{dispersing} or
\textit{active} vs. \textit{inactive} periods of movement \citep{Wilson.2012,
Vasudev.2015}. As a consequence, most permeability surfaces upon which dispersal
corridors are identified are created using GPS relocation data that was
collected on resident individuals. This introduces severe biases and
substantially reduces the power to reveal meaningful corridors, for dispersing
individuals have different needs and motivations compared to resident
individuals (\citeauthor{Killeen.2014}, 2014; \citeauthor{Elliot.2014}, 2014;
\citeauthor{Cozzi.2020}, unpublished). For instance, \cite{Elliot.2014}
demonstrated marked differences in predicted permeability surfaces for
dispersing lions in the KAZA-TFCA depending on whether permeability was assessed
using data collected during residence or dispersal. \cite{Jackson.2016} made
similar observations for African wild dogs and revealed higher permeability for
dispersers than for residents. Moreover, \cite{Abrahms.2017} found a
significantly greater overlap between predicted and observed corridors for three
dispersing wild dogs when predictions were based on information collected only
during wild dog activity, rather than across the entire day.

@@

% Large Study Area
The \textit{extended study area} (centered at -17\degree 13'9''S, 23\degree
56'4''E; elevation ca. 1'030 m, red rectangle in \Cref{StudyArea}a) was confined
by a bounding box encompassing the entire KAZA-TFCA. The KAZA-TFCA lies in the
basins of the Okavango and Zambezi rivers and embraces parts of Angola,
Botswana, Namibia, Zimbabwe, and Zambia. With a total area of over 520,000
km\textsuperscript{2}, it contains several already-existing national parks (NPs)
and other protected areas. It constitutes the earth's largest transboundary
conservation area, which resulted in a rectangular study area of over 1.3 Mio
km\textsuperscript{2}. Similar to the core study area, the extended study area
hosts various landscapes, including savanna, grassland, and dry or moist
woodland habitats. Despite 36 NPs and other protected areas, there is
considerable human influence in some regions of the KAZA-TFCA originating from
farms, high human density, and road traffic.

@@

% Water Layers
Water included open waters, rivers, wetlands, and swamps, whereas anything else
was presumed to be dryland. Because the flood extent of the Okavango Delta is
highly variable within and between years, we created weekly (8 days)
``floodmaps'', following the algorithm developed in \cite{Wolski.2017}. The
algorithm uses thresholding of SWIR reflectances of MODIS Terra satellite
imagery (MCD43A4, \citealp{Schaaf.2015}) to separate water and dryland. As we
rewrote and slightly amended the original Python code in R, we describe the
underlying principles and indicate the discrepancies to the original algorithm
in \Cref{Appendix:FloodmappingAlgorithm}. While water in the Okavango Delta was
mapped dynamically, we used a static watermap for its surroundings. This static
map was based on Globeland's land cover dataset \citep{Chen.2015}, from which we
only retained the categories \textit{wetland} and \textit{water bodies} and
collectively reclassified them to \textit{water}. Globeland had an original
resolution of 30m x 30m, so we coarsened the layer to 250m x 250m using the mode
of each 250m x 250m cell. Lastly, we improved river representation using the
MERIT Hydro dataset \citep{Yamazaki.2019}, from which we added all rivers with a
width of over 10m to our floodmaps.

% Vegetation Layers
The three vegetation layers that we used in addition to the floodmaps were
derived from the MODIS Terra Vegetation Continuous Fields dataset (MOD44B,
\citealp{Dimiceli.2015}). Combined, the three layers added up to 100\% and
depicted the percentage cover of tree-vegetation (henceforth trees),
non-tree-vegetation (henceforth shrubs/grassland), and non-vegetated (henceforth
bare land). To ensure that areas covered by the flood would get 0\% vegetation
cover (i.e. 100\% bare land cover), we used our floodmap that aligned with the
creation date of the layers (30 March 2017) and manually defined pixels covered
by the flood as being 100\% bare land. The MODIS vegetation layers had a
resolution of 250m x 250m, so we did not need to coarsen or interpolate them.

@@

\subsection{Habitat Selection Model}
\label{Modeling}
% Explanation of iSSF and Extraction
We used an integrated step selection function approach (iSSF,
\citealp{Avgar.2016}) to investigate dispersers' selection or avoidance for the
above mentioned spatial covariates. In the iSSF framework, covariates
experienced along realized steps are contrasted with covariates experienced
along alternative random steps that the animal could have taken, but decided not
to. A step is here defined as the connecting line between two consecutive GPS
relocations \citep{Turchin.1998}. In contrast to regular SSFs, iSSFs require to
include movement metrics as covariates in the corresponding conditional logistic
regression model. Their inclusion, in turn, allows simultaneous inference on
habitat and movement preferences, as well as to reduce potential biases in
estimated habitat preferences \citep{Forester.2009, Warton.2013, Avgar.2016}.

To conduct an iSSF analysis, we followed recommendations described in
Appendix S1 of \cite{Avgar.2016}. We prepared our GPS relocation data for
iSSF-analysis using the R-package \textit{amt} \citep{Amt.2019} and coerced
relocations recorded during dispersal to steps that were regularly spaced four
hours apart (allowing for a minor mismatch of up to 15 minutes). Steps that were
separated by more than four hours (e.g. due to GPS failure) were omitted from
further analysis. Each remaining step was paired with 24 random steps, generated
by sampling turning angles from a uniform distribution U(\(-\pi, \pi\)) and step
lengths from a gamma distribution that was fitted using realized step lengths
(shape = 0.3677, scale = 6'302). Together, a realized and its 24 associated
random steps formed a stratum of 25 steps that received a unique identifier.

We extracted spatial covariates along realized and random steps as listed in
\Cref{ExtractedCovars}. For continuous covariates, we calculated the average
value, for binary covariates the percentage cover along the step. Since our
floodmaps were dynamic, we extracted values from the floodmap that was closest
in date to the actual step. We further derived a binary variable indicating
whether or not a step crossed a road and we identified the average distance of
each step to the nearest source of water or the nearest road. We square-rooted
the values for \textit{DistanceToRoads} and \textit{DistanceToWater} to render a
decreasing marginal impact of distance. We standardized extracted covariates to
a mean of zero and a standard deviation of one using a z-score transformation.
Furthermore, we screened covariates for correlation using Pearson's Correlation
Coefficient. None of the covariates were overly correlated (\(|r| > 0.6\),
\citealp{Latham.2011}) and we retained all of them for modeling. To run a proper
iSSF analysis, we also included two movement metrics, namely the cosine of the
turning angle (\(cos(ta)\)) and the logarithm of the step length (\(log(sl)\)),
as covariates in our models \citep{Avgar.2016}. The movement metric \(cos(ta\))
served to describe the directionality of a step, as it transformed the circular
measure of (\(-\pi, +\pi\)) into a linear measure (-1, 1). Thus, positive values
indicated forward movements, whereas negative values indicated backward
movements \citep{Turchin.1998}. The movement metric \(log(sl\)), on the other
hand, was included as an indicator of the preferred step length. Since steps
were regularly spaced by four hours, \(log(sl\)) can also be interpreted as
movement rate.

@@

% Mixed Effects Conditional Logistic Regression Model
To deal with multiple individuals, we applied mixed effects conditional logistic
regression analysis following \cite{Muff.2019}. Their approach allowed to
include individuals for whom some covariates did not vary, as well as to model
random intercepts and slopes for each covariate. We implemented their method
using the R-package \textit{glmmTMB} \citep{Mollie.2017}, where we fixed the
variance of the random intercept to an artificially high value of
10\textsuperscript{12} (for an explanation, see \citealp{Muff.2019}) and used
animal ID to model random intercepts and slopes.

% AIC Model Selection
We defined the movement metrics \(cos(ta)\) and \(log(sl)\) as core covariates
and ran forward model selection based on Akaike's Information Criterion (AIC,
\citealp{Burnham.2002}) for all other spatial covariates. That is, we started
with a basic model and iteratively increased model complexity upon reaching a
full model. We then ranked models according to AIC, assessed relative model
weights, and identified the most parsimonious model. Due to convergence issues,
we were unable to model interactions between spatial covariates.

@@

\subsection{Least-Cost Paths \& Least-Cost Corridors}
To identify movement corridors within the extent of KAZA-TFCA, we calculated
least-cost paths (LCPs) and least-cost corridors (LCCs) between selected points
located in protected areas of our extended study area \citep{Adriaensen.2003,
Sawyer.2011}. Selected points were generated by overlaying the study area with a
regular grid of points, equally spaced 100 km apart and located within protected
areas that contiguously covered at least 700 km\textsuperscript{2}. The
threshold of 700 km\textsuperscript{2} was based on home-range requirements of
African wild dogs reported in \cite{Pomilia.2015} and served to eliminate
locations that were assumed unsuitable for supporting viable wild dog
populations. Finally, we defined centroids as source points for those protected
areas larger than 700 km\textsuperscript{2} that did not get any source points
from the regular grid. In total, we generated 68 points, which resulted in 2'278
unique pairwise combinations and therefore 2'278 unique LCPs and LCCs.

@@

\newpage
\section{Acknowledgements}
All of this work would not have been possible without the aid of several helpful
people and organizations. I would first like to thank Prof. Arpat Ozgul, the
leader and mastermind of our research group, whose expertise in statistics,
biological processes and scientific writing was invaluable for my work.
Moreover, I want to thank Gabriele Cozzi for his excellent guidance, all the
interesting discussions, his constructive criticism, valuable revisions in the
manuscript, and of course his constant supply of freshly brewed coffee.
Particular gratitude goes to Dominik Behr, who supervised my work daily and
greatly contributed to all aspects of the thesis. I genuinely enjoyed the
frequent meetings and brainstorms and I feel privileged to have had the
opportunity of working with such a great scientist. I am also greatly indebted
to Prof. John Fieberg, who visited our department in summer and invested a great
amount of time to consult and improve all statistical aspects of the thesis, as
well as to deepen my understanding of movement ecology. Further, I want to thank
the Okavango Research Institute, especially Piotr Wolski, for sharing all of
their Python scripts, floodmaps, and knowledge on developing dynamic floodmaps.
I also thank all organizations that enabled this project through funds, namely
Basler Stiftung für Biologische Forschung, Claraz Foundation, Idea Wild, Jacot
Foundation, National Geographic Society, Parrotia Stiftung, Stiftung Temperatio,
Wilderness Wildlife Trust Foundation, Forschungskredit der Universität Zürich,
and a Swiss National Foundation Grant (31003A\_182286) to Prof. Arpat Ozgul. I
thank the Ministry of Environment and Tourism of Botswana for granting
permission to conduct this research (Research Permit EWT 8/36/4 XXXVI [33]).
Additionally, I want to thank all of my colleague master students who have been
wonderful companions and made the time at university a blast. Likewise, I thank
my parents for supporting all of my undertakings and for providing first-class
shelter and food. Last, but most definitely not least, I also greatly thank my
girlfriend, who was always there for me and provided welcome distractions for my
mind.

@@

\newpage
\appendix
\section{Appendix}
\subsection{Floodmapping Algorithm}
\label{Appendix:FloodmappingAlgorithm}
To implement the floodmapping algorithm, we defined two sets of polygons located
in the region of the Okavango Delta (see \Cref{Floodmapping}). The first set
consisted of areas known to be permanent dryland, whereas the second set
consisted of permanent waters. Since we were unable to retrieve the original
polygons used in \cite{Wolski.2017}, we geo-referenced one of their
illustrations and traced the polygons there. After recreation of the polygons,
we used the R-package \textit{getSpatialData} \citep{Schwalb.2018} to download
and pre-process all relatively cloud-free MODIS Terra images (MCD43A4,
\citealp{Schaaf.2015}) available for the period of our dispersal events.
Assessment of cloud cover was based on visual inspection of MODIS images on
ORI's website \citep{ORI.2019}. The MCD43A4 MODIS product is particularly useful
for dynamic floodmapping because it is updated with new composite images every
8\textsuperscript{th} day \citep{Wolski.2017}. After download, we classified
each MODIS image into a binary map of water (flood) and dryland using a
threshold that was identified as follows. First, we extracted all reflectance
values of MODIS Terra Band 7 within the water- and dryland-polygons. Second, we
computed histograms of water-reflectances and dryland-reflectances and
empirically verified that reflectances of the two groups were sufficiently
distinct. More specifically, we checked if superimposing the two histograms
resulted in a bimodal plot. This was said to be achieved if the
99\textsuperscript{th} percentile of water-reflectances did not severely exceed
the 1\textsuperscript{st} percentile of dryland-reflectances (\(p_{0.99, water}
- \frac{10}{255} < p_{0.01, dryland}\)). Third, if peaks were sufficiently
different, we calculated a threshold (\(t\)) using \Cref{EQ1}:

@@

\newpage
\subsection{Movement Model \& Dispersal Simulation}
\label{Appendix:DispersalSimulation}
\subsubsection{Methods}
To gain more detailed insights into the movement behavior of dispersing wild
dogs and to simulate dispersal events, we fitted a more complex movement model.
The movement model still contained the most parsimonious habitat selection
model, but in addition included interaction terms between movement metrics (i.e.
\(cos(ta)\), \(log(sl\))) and environmental covariates (e.g. water). For
instance, for the environmental covariate \textit{Water}, we created the
interaction terms \textit{cos(ta):Water} and \textit{log(sl):Water}. These
interactions served to learn how dispersers moved differently, depending on
environmental covariates (for comparable work see \citealp{Prokopenko.2017}). We
applied the same modeling framework as explained in \Cref{Modeling} and ran
forward model selection based on AIC to identify the most parsimonious movement
model. That is, we departed from the habitat selection model and iteratively
added interactions with movement metrics. Note that we applied forward model
selection only on the newly identified interactions, but not for the terms that
were already in the habitat selection model. Since we attempted to keep our
simulation simple, we did not average models with positive AIC-weights but
simply identified the most parsimonious movement model.
